The role of advection versus diffusion in intraventricular transport

### 2025-03-06 paper/main.tex
Results 1

%%--------------RESULTS--------------%%
\section{Results}
\subsection{Motile cilia induce flow compartmentalization within the brain ventricles}

We model the compartmentalized and pulsatile CSF flow observed in vivo~(\Crefpanels{fig:fig1}{b, c})
by the Stokes equations, and solve the equations numerically using a finite element method.
To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we first
consider the simulated CSF velocity and pressure when only modeling
cilia motion, disregarding cardiac pressure pulsations. The
bidirectional traction induced by the motile cilia
inherently leads to a partial compartmentalization of the ventricular
system, with large-scale vortex structures in the anterior and middle
ventricles~(~\Crefpanels{fig:fig1}{e}). The highest velocities ($26.5$ \textmu m/s)
occur in the vicinity of the dorsal cilia in the middle ventricle.

To examine the effect of cardiac pressure pulsations,
we next consider a no-cilia model variation. In this scenario, the vortex structures
generated by the ciliary motion vanish. The simulated CSF
flow is pulsatile as a result of the sinusoidal pressure gradient (~\Crefpanels{fig:fig1}{f, g}),
with a peak velocity magnitude of $4.8$ \textmu m/s. At each time instant, the flow
is unidirectional, with CSF initially flowing rostrocaudally (positive $x-$axis) before changing
direction to caudorostral in the middle of the cardiac cycle.

Finally, we consider the baseline model, simulating flow driven by both cilia
and cardiac pulsations. We observe persistent flow features,
with vortex-structures in the anterior and middle ventricles
and directional flow in the posterior ventricle (~\Crefpanels{fig:fig1}{h--j},
\fixme{Supplementary video 1}). These characteristic patterns suggest
that the flow in the anterior and middle ventricles is majorly driven
by cilia, while the cardiac pressure pulsations dominate
within the posterior ventricle and in the ducts connecting the ventricular
compartments. Over one cardiac cycle, the CSF flow
speed reaches up to 27.7 \textmu m/s, dorsally in the posterior part
of the middle ventricle. This peak is attained at the time instant
coinciding with the time of peak caudal cardiac pulsatile flow, since 
both the tangential traction induced by the cilia and the pulsatile
pressure-driven flow are aligned with the rostrocaudal 
axis. Using the observed maximum velocity magnitude and choosing the height of
the middle ventricle ($110$ \textmu m) as a characteristic length
scale yields a Reynolds number $\mathrm{Re} \approx 0.004$,
justifying the assumption of Stokes flow. In summary, the computational
CSF flow model reproduces the flow features observed
in vivo~\cite{Olstad2019CiliaryDevelopment}.

>>>>

%%--------------RESULTS--------------%%
\section{Results}

\subsection{Motile cilia induce flow compartmentalization within the brain ventricles}

% Why? (Intersting point #1. However, we do not know aspect #2. 
% How? (We therefore ...)
% What ...

In-vivo particle tracking reveals characteristic CSF flow patterns in
zebrafish larval brain ventricles
(\Crefpanels{fig:fig1}{a--c})~\cite{Olstad2019CiliaryDevelopment}. We
first ask whether we can accurately represent these flow patterns with
computational fluid dynamics driven by cardiac pulsations and motile
cilia lining the brain ventricles.

We model the net forces induced by motile cilia as a steady
bidirectional traction acting on parts of the ventricular wall
(\Crefpanels{fig:fig1}{d}). Simulations show that this traction alone
leads to a partial compartmentalization of the ventricular system,
with large-scale vortex structures in the anterior, middle and
posterior ventricles (\Crefpanels{fig:fig1}{e}). The highest flow
speeds ($26.5$ \textmu m/s) occur in the vicinity of the dorsal cilia
in the middle ventricle. To model the cardiac pulsations, we impose a
pulsatile pressure difference between the anterior and posterior
boundaries (\Crefpanels{fig:fig1}{d, \fixme{g}}). The resulting flow
is pulsatile and directional (\Crefpanels{fig:fig1}{\fixme{f}}), and
in the absence of the cilia, the vortex structures vanish. The
velocity magnitude peaks at $4.8$ \textmu m/s. CSF initially flows
rostrocaudally before changing direction to caudorostral in the middle
of the cardiac cycle.

When combining the contributions, we recognize flow features
persisting throughout the cardiac cycle, now with vortex structures in
the anterior and middle ventricles and directional flow in the
posterior ventricle (\Crefpanels{fig:fig1}{h--j}, \fixme{Supplementary
  video 1}). These characteristic patterns suggest that the flow in
the anterior and middle ventricles is primarily driven by the motile
cilia, while the cardiac pressure pulsations dominate within the
posterior ventricle and in the ducts connecting the ventricular
compartments. The CSF flow speed peaks at the time of peak caudal
cardiac pulsatile flow, when both the tangential traction induced by
the cilia and the pulsatile pressure-driven flow are aligned with the
rostrocaudal axis, reaching up to 27.7 \textmu m/s dorsally in the
posterior part of the middle ventricle. The flow Reynolds number is
$\mathrm{Re} \approx 0.004$, justifying the assumption of Stokes
flow. In summary, the computational CSF flow model reproduces the flow
features observed in vivo~\cite{Olstad2019CiliaryDevelopment}.

%% Over one cardiac cycle, the CSF flow speed reaches up to
%% 27.7 \textmu m/s, dorsally in the posterior part of the middle
%% ventricle. This peak is attained at the time instant coinciding with
%% the time of peak caudal cardiac pulsatile flow,

%% Using the observed maximum velocity magnitude and choosing the height
%% of the middle ventricle ($110$ \textmu m) as a characteristic length
%% scale yields a Reynolds number $\mathrm{Re} \approx 0.004$, justifying
%% the assumption of Stokes flow.

%We model the compartmentalized and pulsatile CSF flow observed in vivo~(\Crefpanels{fig:fig1}{b, c}) by the Stokes equations, and solve the equations numerically using a finite element method.

%To characterize CSF flow patterns predicted by
%computational fluid dynamics within the brain ventricles,

%we first consider the simulated CSF velocity and pressure when only
%modeling cilia motion, disregarding cardiac pressure pulsations.
